Emulation of high multiplicity matrix elements for electron-positron and hadron-hadron colliders

### 2022-11-17 qcd.tex
changes to chapter 2

\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Quantum chromodynamics in practice}
\label{chapter:qcd}
    In the previous chapter we introduced the basic concepts
    of QCD and setup the framework of calculating
    cross-sections. Cross-sections are the most relevant
    quantities to compute as they can be measured
    experimentally, and are a property of the particles
    being collided, rather than depending on the
    specifics of the experimental procedure.
    This provides a bridge to compare
    theoretical predictions and experimental measurements.
    The relevant quantity calculated in perturbation theory
    is the partonic cross-section which
    is the matrix element integrated over the relevant
    final-state phase-space, normalised by a flux factor.
    The matrix elements are
    calculated order-by-order in the strong coupling
    $\alpha_{s}$ (c.f. \ref{eqn:matrix_element}),
    where each order corresponds to a set of Feynman diagrams
    that have the appropriate number of vertices,
    as each QCD vertex brings along a factor of $\alpha_{s}$.

    In this chapter we will discuss the problems that
    arise when evaluating these Feynman diagrams and outline
    some solutions that have been widely adopted to circumvent
    these issues in order to provide real-world applicable predictions.
    This chapter forms the theoretical foundations upon which
    this thesis is built on.

    We will consider the divergent nature of matrix
    elements and introduce the most widely used techniques
    to tame these singularities.

\section{Divergent structures}\label{sec:divergences}
    The computation of matrix elements essentially reduces down
    to evaluating Feynman diagrams which are analytical expressions
    built up from the Feynman rules of a theory (c.f. QCD Feynman rules
    Figure \ref{fig:qcd_feynman_rules}). In evaluating certain
    topologies of Feynman diagrams, integral expressions
    containing unconstrained momenta will give rise to
    singularities. The divergences
    associated with high energy modes are ultraviolet
    (UV) divergences, and on the opposite end of the energy
    spectrum, there are infrared (IR) divergences associated
    with low energy modes.

    The de-facto methods to alleviate these
    divergences are through renormalisation
    for UV divergences, and through subtraction
    schemes for IR divergences. Both of these
    methods will be discussed in this section.

@@

    Before we introduce these counterterms, we
    first consider the procedure of regularisation
    which makes these infinities explicitly manifest.
    The most widely adopted method of regularisation
    is dimensional regularisation (DR) \cite{tHooft:1972tcz} which
    is based on the observation that the integral
    (\ref{eqn:loop_integral}), which is carried
    out in $d=4$ space-time dimensions, would be finite
    if we move away from $d=4$ dimensions
    \footnote{In dimensional regularisation it is possible to regulate
    the integral regardless of if $d > 4$ or $d < 4$, it would amount
    to a sign difference.}.

    In DR we define $d = 4 - 2\epsilon$ such that
    after Feynman parametrisation (\ref{eqn:loop_integral})
    becomes
    \begin{equation}\label{eqn:dim_reg}
        I_{\mathrm{UV}} = \int \dfrac{\mathrm{d}^{d}\ell}{(2\pi)^{d}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} = \dfrac{\mathrm{i}}{(4\pi)^{2}}\left(\dfrac{-p^{2}}{4\pi}\right)^{-\epsilon}\dfrac{\Gamma(1-\epsilon)^{2}}{\Gamma(2-2\epsilon)} \, \Gamma(\epsilon) \, ,
    \end{equation}
    where the divergence for $d = 4$, or equivalently
    $\epsilon \rightarrow 0$, is now captured in the
    Gamma function, $\Gamma(\epsilon)$. This becomes clear once
    we expand $\Gamma(\epsilon)$ around small $\epsilon$
    \begin{equation}\label{eqn:gamma_laurent}
        \Gamma(\epsilon) = \dfrac{1}{\epsilon} - \gamma_{E} + \dfrac{1}{2}\left(\gamma_{E}^{2} + \dfrac{\pi^{2}}{6}\right)\epsilon + \mathcal{O}(\epsilon^{2}) \, ,
    \end{equation}
    where $\gamma_{E} \approx 0.577$ is the Euler-Mascheroni
    constant, meaning the divergence is now regularised as a pole in
    $\epsilon$.

@@

    A subtlety with adjusting the dimension of the theory
    from $d=4$ to arbitrary dimensions is that the mass dimension
    of the Lagrangian has to change accordingly to retain a
    dimensionless action in natural units. We can account for this by
    introducing an arbitrary energy scale $\mu_{R}$, the renormalisation
    scale, into the gauge coupling
    constant to modify the mass dimension of the Lagrangian.
    For QCD this is done via the modification
    \begin{equation}\label{eqn:dimensionful_coupling}
        g_{s} \rightarrow g_{s}\mu_{R}^{\epsilon} \, .
    \end{equation}
    where it is understood that the original gauge coupling is
    dimensionless. The order of $\mu_{R}$ is determined by examining
    the dimensions of the gauge and fermion fields
    in the Lagrangian (\ref{eqn:L_QCD}).
    We see that this introduces a scale dependence
    on the coupling constant $\alpha_{s}$, a point we will return
    to in Section~\ref{sec:alpha_running}.
    % In a calculation,
    % intermediate steps will have a dependence on $\mu$,
    % however, once the $\epsilon \rightarrow 0$ limit
    % is taken the scale dependence drops out.

    The inclusion of counterterms into the theory can be
    thought of as redefinitions of the bare fields
    and couplings in the QCD Lagrangian (\ref{eqn:L_QCD}) to restore predictive
    power to the theory. The renormalised fields and couplings
    can be given as (see for instance Ref.~\cite{mangano1999introduction})
    \begin{align}\label{eqn:renormlised_fields}
        \psi_{\mathrm{bare}} &= \sqrt{Z_{2}} \psi_{R} \, , \nonumber \\
        A_{\mathrm{bare}}^{\mu} &= \sqrt{Z_{3}} A_{R}^{\mu} \, ,\\
        g_{s,\mathrm{bare}} &= Z_{g} \mu_{R}^{\epsilon} g_{s,R} \, , \nonumber
    \end{align}
    where it conventional to define $Z_{1} = Z_{g}Z_{2}\sqrt{Z_{3}}$
    such that we can set $Z_{n} = 1 + \delta_{n}$ for $n \in \{1, 2, 3\}$.
    In this way we can write the Lagrangian as
    \begin{equation}
        \mathcal{L}_{R} = \mathcal{L}_{\mathrm{bare}} + \mathcal{L}_{\mathrm{c.t.}} \, ,
    \end{equation}
    where the counterterms appearing in $\mathcal{L}_{\mathrm{c.t.}}$
    are determined by calculating $\delta_{n}$. There is freedom
    in the choice of $\delta_{n}$ which gives rise to
    different regularisation schemes. The minimal subtraction
    (MS) scheme subtracts only the epsilon pole appearing
    in loop integrals. However, the most
    commonly used regularisation scheme is the modified minimal
    subtraction ($\overline{\mathrm{MS}}$) scheme which subtracts the
    epsilon pole plus a universal constant appearing in all
    loop integrals.

    In an all orders calculation of an observable,
    there would be no dependence on the renormalisation
    scale as it is a remnant of the regularisation prescription
    However, because in perturbative QCD observables
    are calculated order-by-order, there will be a residual
    dependence on the renormalisation scale stemming from
    the missing higher order terms. The conventional way
    that the community estimates the uncertainty associated
    with this residual dependence is by carrying out
    a scale variation (see Section~\ref{sec:scale_variations}).

@@

    The renormalisation scale dependence of $\alpha_{s}$
    is governed by the Callan-Symanzik \cite{Callan:1970yg,Symanzik:1970rt}
    $\beta$-function
    \begin{equation}\label{eqn:beta_fn}
        \mu_{R}^{2} \dfrac{\partial \alpha_{s}(\mu_{R}^{2})}{\partial \mu_{R}^{2}} = \beta(\alpha_{s}) \, ,
    \end{equation}
    where the $\beta$-function can be written as a perturbative
    expansion in $\alpha_{s}$
    \begin{equation}\label{eqn:beta_series}
        -\beta(\alpha_{s}) = \alpha_{s}\sum_{n=0}^{\infty}\left(\dfrac{\alpha_{s}}{4\pi}\right)^{n+1} \beta_{n}\, 
    \end{equation}
    where the coefficients $\beta_{n}$ have been computed
    up to $\beta_{4}$ \cite{Baikov:2016tgj,Luthe:2017ttg}.
    The solution to (\ref{eqn:beta_fn}) to first order
    is
    \begin{equation}\label{eqn:1l_alpha}
        \alpha_{s}(\mu_{R}^{2}) = \dfrac{1}{\frac{\beta_{0}}{4\pi}\log\left(\frac{\mu_{R}^{2}}{\Lambda_{\mathrm{QCD}}^{2}}\right)} \, ,
    \end{equation}
    where $\Lambda_{\mathrm{QCD}}$ is the QCD scale,
    the scale at which $\alpha_{s}$ becomes
    large enough that perturbation theory breaks down,
    and $\beta_{0} = (11C_{A} - 4T_{R}n_{f})/3$.
    In QCD, $C_{A}=3$ and $T_{R}=\frac{1}{2}$, meaning for
    $n_{f} < \frac{33}{2}$ the sign of $\beta_{0}$ is positive.
    In nature we have observed six flavours of quarks, so
    $\beta_{0} > 0$. In fact all $\beta$-coefficients
    computed to date are positive, meaning that $\alpha_{s}$
    decreases with increasing energy due to the minus sign
    in (\ref{eqn:beta_series}). This property is known as
    asymptotic freedom \cite{Gross:1973id,Politzer:1973fx},
    and is justification for treating QCD perturbatively
    when the energy scale is high, such as at collider experiments.
    The running of the coupling constant has been
    observed experimentally as illustrated in Figure~\ref{fig:alpha_s_running}.

    \begin{figure}
        \includegraphics{qcd/alpha_s_running.pdf}
        \caption{The running of the strong coupling constant $\alpha_{s}$,
        as determined by experiments, with QCD theory prediction in black.
        Figure reference \cite{Workman:2022ynf}.}
        \label{fig:alpha_s_running}
    \end{figure}

@@

\subsection{Infrared divergences}\label{sec:IR_divergences}
    In Section~\ref{sec:UV} we saw how divergences
    stemming from high energy behaviour would occur
    in loop diagrams. On the other end of the energy
    spectrum we also have divergences from low energy
    modes. Since these divergences emerge from low energy
    behaviour they are called infrared (IR) divergences. There
    are two cases in which these divergences arise:
    \begin{itemize}
        \item virtual divergences in loop integrals,
        \item real emission divergences when an emission
    of an extra particle has vanishing energy (soft),
    or becomes parallel to an external leg (collinear).
    \end{itemize}

    Virtual IR divergences can be understood by looking
    at an integral of the form
    \begin{equation}\label{eqn:virtual_IR_integral}
        I_{\mathrm{V}} = \int \dfrac{\mathrm{d}^{4}\ell}{(2\pi)^{4}} \dfrac{1}{\ell^{2}(\ell+p_{1})^{2}(\ell-p_{2})^{2}} \, ,
    \end{equation}
    which is encountered when calculating the one-loop
    virtual correction to the gluon-quark-antiquark vertex
    as shown in Figure~\ref{fig:vertex_correction}.
    It is clear that the denominator vanishes when $\ell \rightarrow 0$
    or when either $(\ell + p_{1})^{2}$ or $(\ell - p_{2})^{2} \rightarrow 0$.
    These situations correspond to the gluon in the loop propagator
    going soft, or collinear to the external quark or antiquark.
    These divergences are regulated in DR to give
    \begin{equation}\label{eqn:virtual_IR_DR}
        I_{\mathrm{V}} = \dfrac{\mathrm{i}}{(4\pi)^{2}Q^{2}}\left(\dfrac{-Q^{2}}{4\pi}\right)^{-\epsilon}\dfrac{\Gamma(1+\epsilon)\Gamma(1-\epsilon)^{2}}{\Gamma(1-2\epsilon)}\left[\dfrac{1}{\epsilon^{2}}\right] \, ,
    \end{equation}
    where $Q = (p_{1} + p_{2})$. There is a double $\epsilon$ pole manifest,
    corresponding to the associated soft and collinear divergences in real emission corrections.

@@

    Real IR divergences are manifest when integrating
    matrix elements over the phase-space of external state momenta.
    Matrix elements can be written as functions of Mandelstam variables
    \begin{equation}\label{eqn:mandelstam}
        s_{ij} = (p_{i} + p_{j})^{2} \xrightarrow[\mathrm{limit}]{\mathrm{massless}} 2 E_{i}E_{j}(1-\cos \theta_{ij}) \, ,
    \end{equation}
    where $p_{i}$ and $p_{j}$ are the 4-momenta of partons $i$ and $j$
    in the hard scattering.
    When either of the partons become soft, $E_{i,j}\rightarrow 0$,
    or when they go collinear to each other, $\theta_{ij} \rightarrow 0$,
    the matrix element will diverge if $s_{ij}$ appears
    in the denominator. These scenarios correspond to an external
    particle becoming unresolved.

@@

    At a detector of a collider experiment, there is a finite
    energy resolution of the calorimeter. This means that
    it is not physically possible to observe arbitrarily soft
    partons. It is also not possible to distinguish two partons
    at arbitrarily small angles from one parton with the combined
    momenta. These limitations along with the KLN theorem
    means that any physical observable that we wish to predict
    with fixed-order perturbation theory must be insensitive to the
    emissions of soft or collinear partons. Any observable obeying
    this criteria is an infrared and collinear (IRC) safe observable.
    This leads to the concept of jets \cite{Salam:2010nqg} which are collimated
    particles combined according to some jet algorithm.

    In DR, the singularities from soft and collinear divergences
    are manifest as $\epsilon$ poles which makes the cancellation
    simple between the real and virtual parts. However, in practice, 
    the phase-space integrals of the matrix elements are rarely carried
    out analytically due to the high dimensionality of the integral,
    rendering them intractable.
    Instead they are done numerically through the use
    of Monte Carlo methods (see Chapter~\ref{chap:MC}),
    which requires the integral to be in
    integer dimensions. Therefore, numerical techniques
    have been developed to deal with these IR divergences as well. The
    most common technique, subtraction, will be discussed in
    more detail in Section~\ref{sec:subtraction}.

@@

    We will inspect the soft and collinear limits separately
    to see how the matrix element factorises in these respective
    limits.
    The following section will follow the conventions
    of Ref.~\cite{Catani:1996vz}, namely that the colour and
    helicity summed $n$-body matrix element written as
    \begin{equation}\label{eqn:CS_matrix_element}
        |\mathcal{M}_{n}|^{2} = {}_{n}\langle 1, \ldots, n | 1, \ldots, n \rangle_{n}
    \end{equation}
    where $| 1, \ldots, n \rangle_{n}$ is a vector
    in colour + helicity space (see Appendix~\ref{appendix:catani_seymour}
    for a more thorough explanation on notation).

\subsection{Soft limits}\label{sec:me_soft}
    Consider a tree-level matrix element $|\mathcal{M}_{n+1}|^{2}$
    with a final-state gluon $j$. Note that reduced matrix
    elements associated with taking a quark soft have to vanish
    due to violation of quark number, therefore only gluons
    are considered.
    The limit of the soft gluon with momentum
    $p_{j}$ can be parametrised by
    \begin{equation}\label{eqn:soft_parametrisation}
        p_{j}^{\mu} = \lambda q^{\mu} \, , \quad \lambda \rightarrow 0 \, ,
    \end{equation}
    where $q^{\mu}$ is an arbitrary four-vector and $\lambda$
    is a scale parameter. In this limit the matrix element
    can be written as
    \begin{equation}\label{eqn:soft_factorisation}
        \begin{split}
        |\mathcal{M}_{n+1}|^{2} \rightarrow &-\dfrac{1}{\lambda^{2}}8\pi \mu^{2\epsilon}\alpha_{s} \sum_{i}\dfrac{1}{p_{i}q} \sum_{k \neq i} \dfrac{p_{k}p_{i}}{(p_{i} + p_{k})q} \\
        &{}_{n}\langle 1, \ldots, j-1, j+1, \ldots, n+1 | \bm{T}_{k} \cdot \bm{T}_{i} | 1, \ldots, j-1, j+1, \ldots, n+1 \rangle_{n} \, ,
        \end{split}
    \end{equation}
    where terms less singular than $1 / \lambda^{2}$ have been
    neglected. The reduced matrix element represented by the
    colour + helicity vector inner product is obtained by removing
    the soft gluon from the $(n+1)$-body matrix element.
    The indices $i$, $j$, and $k$ label partons involved
    in the factorisation process: $i$ is the emitter parton,
    $j$ is the emitted parton, and $k$ is a parton accounting
    for the colour correlations. The scale $\mu$ can be identified as the
    renormalisation scale, and $\bm{T_{i}}$ is the colour-charge
    operator. It acts on the colour vectors in the reduced matrix element
    to give factors of $C_{A}$ and $C_{F}$, hence this factorisation is not exact.

    The matrix elements in (\ref{eqn:soft_factorisation}) are
    unambiguously defined only when momentum conservation is
    fulfilled. This is only true in the strict $\lambda = 0$
    limit. Away from the limit, care has to be taken to conserve
    momentum conservation and keep the relevant partons on-shell.
    This can be done through the use of momentum mappings
    \cite{Catani:1996vz,Kosower:1997zr}.

@@

    The analogous expression of (\ref{eqn:F_collinear_factorisation})
    for initial-state splitting is
    \begin{equation}\label{eqn:I_collinear_factorisation}
        |\mathcal{M}_{n+1}|^{2} \rightarrow \dfrac{1}{x} \dfrac{1}{p_{i}p_{a}} 4\pi \mu^{2\epsilon} \alpha_{s} \; {}_{n} \langle 1,\ldots,n+1;ai, \ldots | \hat{P}_{ai}(x, k_{\perp}) | 1,\ldots,n+1;ai,\ldots \rangle_{n} \, ,
    \end{equation}
    where the replacement of partons $a$ and $i$ have been made
    explicit by the presence of the parton $ai$ in the colour + helicity vector.
    The specific type of parton $ai$ depends upon the types of $a$ and $i$.

>>>>

\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Quantum chromodynamics in practice}
\label{chapter:qcd}
    In the previous chapter we introduced the basic concepts
    of QCD and setup the framework of calculating
    cross-sections. Cross-sections are the most relevant
    quantities to compute as they can be measured
    experimentally, and are a property of the particles
    being collided, rather than depending on the
    specifics of the experimental procedure.
    This provides a bridge to compare
    theoretical predictions and experimental measurements.
    The relevant quantity calculated in perturbation theory
    is the partonic cross-section which
    is the matrix element integrated over the relevant
    final-state phase-space, normalised by a flux factor.
    The matrix elements are
    calculated order-by-order in the strong coupling
    $\alpha_{s}$ (c.f. Eq. \ref{eqn:matrix_element}),
    where each order corresponds to a set of Feynman diagrams
    that have the appropriate number of vertices,
    as each QCD vertex brings along a factor of $\alpha_{s}$.

    In this chapter we will discuss the problems that
    arise when evaluating these Feynman diagrams and outline
    some solutions that have been widely adopted to circumvent
    these issues in order to provide real-world applicable predictions.
    We will consider the divergent nature of matrix
    elements and introduce the most widely used techniques
    to tame these singularities.
    This chapter forms the theoretical foundations upon which
    this thesis is built on.

\section{Divergent structures}\label{sec:divergences}
    The computation of matrix elements essentially reduces
    to evaluating Feynman diagrams which are analytical expressions
    built up from the Feynman rules of a theory (c.f. QCD Feynman rules
    Figure \ref{fig:qcd_feynman_rules}). In evaluating certain
    topologies of Feynman diagrams, integral expressions
    containing unconstrained momenta will give rise to
    singularities. The divergences
    associated with high energy modes are ultraviolet
    (UV) divergences, and on the opposite end of the energy
    spectrum, there are infrared (IR) divergences associated
    with low energy modes.

    The de facto methods to alleviate these
    divergences are through renormalisation
    for UV divergences, and through subtraction
    schemes for IR divergences. Both of these
    methods will be discussed in this section.

@@

    Before we introduce these counterterms, we
    first consider the procedure of regularisation
    which makes these infinities explicitly manifest.
    The most widely adopted method of regularisation
    is dimensional regularisation (DR) \cite{tHooft:1972tcz} which
    is based on the observation that the integral
    in Eq. (\ref{eqn:loop_integral}), which is carried
    out in $d=4$ space-time dimensions, would be finite
    if we move away from $d=4$ dimensions\footnote{In dimensional regularisation it is possible to regulate
    the integral regardless of if $d > 4$ or $d < 4$, it would amount
    to a sign difference.}.

    In DR we define $d = 4 - 2\epsilon$ such that
    after Feynman parametrisation Eq. (\ref{eqn:loop_integral})
    becomes
    \begin{equation}\label{eqn:dim_reg}
        I_{\mathrm{UV}} = \int \dfrac{\mathrm{d}^{d}\ell}{(2\pi)^{d}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} = \dfrac{\mathrm{i}}{(4\pi)^{2}}\left(\dfrac{-p^{2}}{4\pi}\right)^{-\epsilon}\dfrac{\Gamma(1-\epsilon)^{2}}{\Gamma(2-2\epsilon)} \, \Gamma(\epsilon) \, ,
    \end{equation}
    where the divergence for $d = 4$, or equivalently
    $\epsilon \rightarrow 0$, is now captured in the
    Gamma function, $\Gamma(\epsilon)$. This becomes clear once
    we expand $\Gamma(\epsilon)$ around small $\epsilon$
    \begin{equation}\label{eqn:gamma_laurent}
        \Gamma(\epsilon) = \dfrac{1}{\epsilon} - \gamma_{E} + \dfrac{1}{2}\left(\gamma_{E}^{2} + \dfrac{\pi^{2}}{6}\right)\epsilon + \mathcal{O}(\epsilon^{2}) \, ,
    \end{equation}
    where $\gamma_{E} \approx 0.577$ is the Euler-Mascheroni
    constant, meaning the divergence is now regularised as a pole in
    $\epsilon$.

@@

    A subtlety with adjusting the dimension of the theory
    from $d=4$ to arbitrary dimensions is that the mass dimension
    of the Lagrangian has to change accordingly to retain a
    dimensionless action in natural units. We can account for this by
    introducing an arbitrary energy scale $\mu_{R}$, the renormalisation
    scale, into the gauge coupling
    constant to modify the mass dimension of the Lagrangian.
    For QCD this is done via the modification
    \begin{equation}\label{eqn:dimensionful_coupling}
        g_{s} \rightarrow g_{s}\mu_{R}^{\epsilon} \, .
    \end{equation}
    where it is understood that the original gauge coupling is
    dimensionless. The order of $\mu_{R}$ is determined by examining
    the dimensions of the gauge and fermion fields
    in the Lagrangian in Eq. (\ref{eqn:L_QCD}).
    We see that this introduces a scale dependence
    on the coupling constant $\alpha_{s}$, a point we will return
    to in Section~\ref{sec:alpha_running}.
    % In a calculation,
    % intermediate steps will have a dependence on $\mu$,
    % however, once the $\epsilon \rightarrow 0$ limit
    % is taken the scale dependence drops out.

    The inclusion of counterterms into the theory can be
    thought of as redefinitions of the bare fields
    and couplings in the QCD Lagrangian to restore predictive
    power to the theory. The renormalised fields and couplings
    can be given as (see for instance Ref.~\cite{mangano1999introduction})
    \begin{align}\label{eqn:renormlised_fields}
        \psi_{\mathrm{bare}} &= \sqrt{Z_{2}} \psi_{R} \, , \nonumber \\
        A_{\mathrm{bare}}^{\mu} &= \sqrt{Z_{3}} A_{R}^{\mu} \, ,\\
        g_{s,\mathrm{bare}} &= Z_{g} \mu_{R}^{\epsilon} g_{s,R} \, , \nonumber
    \end{align}
    where it conventional to define $Z_{1} = Z_{g}Z_{2}\sqrt{Z_{3}}$
    such that we can set $Z_{n} = 1 + \delta_{n}$ for $n \in \{1, 2, 3\}$.
    In this way we can write the renormalised Lagrangian as
    \begin{equation}
        \mathcal{L}_{\mathrm{renorm}} = \mathcal{L}_{\mathrm{bare}} + \mathcal{L}_{\mathrm{c.t.}} \, ,
    \end{equation}
    where the counterterms appearing in $\mathcal{L}_{\mathrm{c.t.}}$
    are determined by calculating $\delta_{n}$. There is freedom
    in the choice of $\delta_{n}$ which gives rise to
    different regularisation schemes. The minimal subtraction
    (MS) scheme subtracts only the epsilon pole appearing
    in loop integrals. However, the most
    commonly used regularisation scheme is the modified minimal
    subtraction ($\overline{\mathrm{MS}}$) scheme which subtracts the
    epsilon pole along with a universal constant appearing in all
    loop integrals.

    In an all orders calculation of an observable,
    there would be no dependence on the renormalisation
    scale as it is a remnant of the regularisation prescription.
    However, because in perturbative QCD observables
    are calculated order-by-order, there will be a residual
    dependence on the renormalisation scale stemming from
    the missing higher order terms. The conventional way
    that theoretical uncertainties associated
    with this residual dependence are determined is by carrying out
    a scale variation (see Section~\ref{sec:scale_variations}).

@@

    The renormalisation scale dependence of $\alpha_{s}$
    is governed by the Callan-Symanzik 
    $\beta$-function \cite{Callan:1970yg,Symanzik:1970rt}
    \begin{equation}\label{eqn:beta_fn}
        \mu_{R}^{2} \dfrac{\partial \alpha_{s}(\mu_{R}^{2})}{\partial \mu_{R}^{2}} = \beta(\alpha_{s}) \, ,
    \end{equation}
    where the $\beta$-function can be written as a perturbative
    expansion in $\alpha_{s}$
    \begin{equation}\label{eqn:beta_series}
        -\beta(\alpha_{s}) = \alpha_{s}\sum_{n=0}^{\infty}\left(\dfrac{\alpha_{s}}{4\pi}\right)^{n+1} \beta_{n}\, 
    \end{equation}
    where the coefficients $\beta_{n}$ have been computed
    up to $\beta_{4}$ \cite{Baikov:2016tgj,Luthe:2017ttg}.
    The solution to Eq. (\ref{eqn:beta_fn}) to first order
    is
    \begin{equation}\label{eqn:1l_alpha}
        \alpha_{s}(\mu_{R}^{2}) = \dfrac{1}{\frac{\beta_{0}}{4\pi}\log\left(\frac{\mu_{R}^{2}}{\Lambda_{\mathrm{QCD}}^{2}}\right)} \, ,
    \end{equation}
    where $\beta_{0} = (11C_{A} - 4T_{R}n_{f})/3$.
    In QCD, $C_{A}=3$ and $T_{R}=\frac{1}{2}$, meaning for
    $n_{f} < \frac{33}{2}$ the sign of $\beta_{0}$ is positive.
    In nature we have observed six flavours of quarks, so
    $\beta_{0} > 0$. In fact all $\beta$-coefficients
    computed to date are positive, meaning that $\alpha_{s}$
    decreases with increasing energy due to the minus sign
    in Eq. (\ref{eqn:beta_series}). This property is known as
    asymptotic freedom \cite{Gross:1973id,Politzer:1973fx},
    and is justification for treating QCD perturbatively
    when the energy scale is high, such as at collider experiments.
    The running of the coupling constant has been
    observed experimentally as illustrated in Figure~\ref{fig:alpha_s_running}.

    \begin{figure}
        \includegraphics{qcd/alpha_s_running.pdf}
        \caption{The running of the strong coupling constant $\alpha_{s}$,
        as determined by experiments, with QCD theory prediction in black.
        Figure reproduced from \cite{Workman:2022ynf}.}
        \label{fig:alpha_s_running}
    \end{figure}

@@

\subsection{Infrared divergences}\label{sec:IR_divergences}
    In Section~\ref{sec:UV} we saw how divergences
    stemming from high energy behaviour would occur
    in loop diagrams. On the other end of the energy
    spectrum we also have divergences from low energy
    modes. Since these divergences emerge from low energy
    behaviour they are called infrared divergences. There
    are two cases in which these divergences arise:
    \begin{itemize}
        \item virtual divergences in loop integrals,
        \item real-emission divergences when an emission
    of an extra particle has vanishing energy (soft),
    or becomes parallel to an external leg (collinear).
    \end{itemize}

    Virtual IR divergences can be understood by looking
    at an integral of the form
    \begin{equation}\label{eqn:virtual_IR_integral}
        I_{\mathrm{V}} = \int \dfrac{\mathrm{d}^{4}\ell}{(2\pi)^{4}} \dfrac{1}{\ell^{2}(\ell+p_{1})^{2}(\ell-p_{2})^{2}} \, ,
    \end{equation}
    which is encountered when calculating the one-loop
    virtual correction to the gluon-fermion vertex
    as shown in Figure~\ref{fig:vertex_correction}.
    It is clear that the denominator vanishes when $\ell \rightarrow 0$
    or when either $(\ell + p_{1})^{2}$ or $(\ell - p_{2})^{2} \rightarrow 0$.
    These situations correspond to the gluon in the loop propagator
    going soft, or collinear to the external quarks, respectively.
    These divergences are regulated in DR to give
    \begin{equation}\label{eqn:virtual_IR_DR}
        I_{\mathrm{V}} = \dfrac{\mathrm{i}}{(4\pi)^{2}Q^{2}}\left(\dfrac{-Q^{2}}{4\pi}\right)^{-\epsilon}\dfrac{\Gamma(1+\epsilon)\Gamma(1-\epsilon)^{2}}{\Gamma(1-2\epsilon)}\left[\dfrac{1}{\epsilon^{2}}\right] \, ,
    \end{equation}
    where $Q = (p_{1} + p_{2})$. In this expression, there is a double $\epsilon$ pole manifest,
    corresponding to the associated soft and collinear divergences in real emission corrections.

@@

    Real IR divergences arise when integrating
    matrix elements over the phase-space of external state momenta.
    Matrix elements can be written as functions of Mandelstam variables
    \begin{equation}\label{eqn:mandelstam}
        s_{ij} = (p_{i} + p_{j})^{2} \xrightarrow[\mathrm{limit}]{\mathrm{massless}} 2 E_{i}E_{j}(1-\cos \theta_{ij}) \, ,
    \end{equation}
    where $p_{i}$ and $p_{j}$ are the 4-momenta of two partons $i$ and $j$
    in the hard scattering.
    When either of the partons become soft, $E_{i,j}\rightarrow 0$,
    or when they go collinear to each other, $\theta_{ij} \rightarrow 0$,
    the matrix element will diverge if $s_{ij}$ appears
    in the denominator. These scenarios correspond to an external
    particle becoming unresolved.

@@

    At a detector of a collider experiment, there is a finite
    energy resolution of the calorimeter. This means that
    it is not physically possible to observe arbitrarily soft
    particles. It is also not possible to distinguish two particles
    at arbitrarily small angles from one particle with the combined
    momenta. These limitations along with the KLN theorem
    means that any physical observable that we wish to predict
    with fixed-order perturbation theory must be insensitive to the
    emissions of soft or collinear particles. Any observable obeying
    this criteria is an infrared and collinear (IRC) safe observable.
    This leads to the concept of jets \cite{Salam:2010nqg} which are collimated
    particles combined according to some jet algorithm.

    In DR, the singularities from soft and collinear divergences
    are manifest as $\epsilon$ poles which makes the cancellation
    simple between the real and virtual parts. However, in practice, 
    the phase-space integrals of the matrix elements are rarely carried
    out analytically due to the high dimensionality of the integral,
    rendering them intractable.
    Instead they are done numerically through the use
    of Monte Carlo methods (see Chapter~\ref{chapter:MC}),
    which requires the integral to be in
    integer dimensions. Therefore, numerical techniques
    have been developed to deal with these IR divergences as well. The
    most common technique, subtraction, will be discussed in
    more detail in Section~\ref{sec:subtraction}.

@@

    We will inspect the soft and collinear limits separately
    to see how the matrix element factorises in these respective
    limits.
    The following section will follow the conventions
    of Ref.~\cite{Catani:1996vz}, namely that the colour and
    helicity summed $n$-body matrix element can be written as
    \begin{equation}\label{eqn:CS_matrix_element}
        |\mathcal{M}_{n}|^{2} = {}_{n}\langle 1, \ldots, n | 1, \ldots, n \rangle_{n}
    \end{equation}
    where $| 1, \ldots, n \rangle_{n}$ is a vector
    in colour + helicity space (see Appendix~\ref{appendix:catani_seymour}
    for a more thorough explanation on notation).

\subsection{Soft limits}\label{sec:me_soft}
    Consider a tree-level matrix element $|\mathcal{M}_{n+1}|^{2}$
    with a final-state gluon $j$. Note that reduced matrix
    elements associated with taking a quark soft have to vanish
    due to violation of quark number, therefore only gluons
    are considered.
    The limit of the soft gluon with momentum
    $p_{j}$ can be parametrised as
    \begin{equation}\label{eqn:soft_parametrisation}
        p_{j}^{\mu} = \lambda q^{\mu} \, , \quad \lambda \rightarrow 0 \, ,
    \end{equation}
    where $q^{\mu}$ is an arbitrary four-vector and $\lambda$
    is a scale parameter. In this limit the matrix element
    can be written as
    \begin{equation}\label{eqn:soft_factorisation}
        \begin{split}
        |\mathcal{M}_{n+1}|^{2} \rightarrow &-\dfrac{1}{\lambda^{2}}8\pi \mu^{2\epsilon}\alpha_{s} \sum_{i}\dfrac{1}{p_{i}q} \sum_{k \neq i} \dfrac{p_{k}p_{i}}{(p_{i} + p_{k})q} \\
        &{}_{n}\langle 1, \ldots, j-1, j+1, \ldots, n+1 | \bm{T}_{k} \cdot \bm{T}_{i} | 1, \ldots, j-1, j+1, \ldots, n+1 \rangle_{n} \, ,
        \end{split}
    \end{equation}
    where terms less singular than $1 / \lambda^{2}$ have been
    neglected. The reduced matrix element represented by the
    colour + helicity vector inner product is obtained by removing
    the soft gluon, $j$, from the $(n+1)$-body matrix element.
    The indices $i$, $j$, and $k$ label partons involved
    in the factorisation process: $i$ is the emitter parton,
    $j$ is the emitted parton, and $k$ is a parton accounting
    for the colour correlations.
    The scale $\mu$ can be identified as the
    renormalisation scale, and $\bm{T}_{i}$ is the colour-charge
    operator. It acts on the colour vectors in the reduced matrix element
    to give factors of $C_{A}$ and $C_{F}$, hence this factorisation is not exact.

    The matrix elements in Eq. (\ref{eqn:soft_factorisation}) are
    unambiguously defined only when momentum conservation is
    fulfilled. This is only true in the strict $\lambda = 0$
    limit. Away from the limit, care has to be taken to conserve
    momentum conservation and keep the relevant partons on-shell.
    This can be done through the use of momentum mappings
    \cite{Catani:1996vz,Kosower:1997zr}.

@@

    The analogous expression of Eq. (\ref{eqn:F_collinear_factorisation})
    for initial-state splitting is
    \begin{equation}\label{eqn:I_collinear_factorisation}
        |\mathcal{M}_{n+1}|^{2} \rightarrow \dfrac{1}{x} \dfrac{1}{p_{i}p_{a}} 4\pi \mu^{2\epsilon} \alpha_{s} \; {}_{n} \langle 1,\ldots,n+1;ai, \ldots | \hat{P}_{ai}(x, k_{\perp}) | 1,\ldots,n+1;ai,\ldots \rangle_{n} \, ,
    \end{equation}
    where the replacement of partons $a$ and $i$ have been made
    explicit by the presence of the parton $ai$ in the colour + helicity vectors.
    The specific type of parton $ai$ depends upon the types of $a$ and $i$.

    Similar to Eq. (\ref{eqn:soft_factorisation}) where the factorisation was
    only true in the strict soft limit, Eqns. (\ref{eqn:F_collinear_factorisation})
    and (\ref{eqn:I_collinear_factorisation}) are only
    true in the strict collinear limit. Away from these
    limits, care has to be taken to conserve momenta
    via the use of momenta mappings.
